---
title: viewscape in R
author: Xiaohao Yang
date: '2024-02-13'
slug: viewscape
categories: ["tutorials"]
tags: ["R", "R package"]
---

This is an integrated tutorial about `viewscape` and `dsmSearch`. These two packages were initially one before getting published because I thought it would be cool if one can just specify a geographical area without DSM/DEM to compute a viewshed(s). This has been explained the purpose of making the `dsmSearch`, which was two functions for searching and downloading point cloud data via API.

Of course, `viewscape` is not designed for computing the viewshed even if this is the core of the package. A set of function for calculating viewscape metrics ([see details](https://github.com/land-info-lab/viewscape/blob/master/README.md)) is the main contribution of this package. What's more, it provides the functionality to subset the viewshed based on the angle and orientation of the field of view. 

What is 'viewshed' and what is 'viewscape' by the way? Basically, I would say viewscape is a concept of analysis relying on viewshed, which is the visible area that can be seen from the  location of a observer. Viewscape analysis is to understand the composition of landscape and the structure of viewshed itself and visible objects within the viewshed.

```{r eval = FALSE}
# install packages
install.pakcages("lidR")
install.pakcages("viewscape")
install.pakcages("dsmSearch")
```

### 1.Compute viewshed
To compute a viewshed, a viewpoint and a DSM are essential. Let's download the LiDAR data from a small part of a city. In this tutorial, I take the Central Park (NYC) as an example. First, a bounding box (bbox) is used to locate the area to download the data. `lidR`, a powerful tool for processing LiDAR data, is used to generate DSM/DEM.

```{r eval = FALSE}
# bbox finder: http://bboxfinder.com/
bbox <- c(-73.976440,40.768793,-73.969628,40.773295)

# download LiDAR (.las) 
las <- dsmSearch::get_lidar(bbox = bbox, epsg = 2263)
# constrct DSM and DEM using .las data
dsm <- lidR::rasterize_canopy(las, 5, lidR::dsmtin())
dem <- lidR::rasterize_terrain(las, 5, lidR::tin())
```

In this case, I just set the center point of the DSM as the viewpoint. 

```{r eval = FALSE}
row <- trunc(terra::nrow(dsm)/2)
col <- trunc(terra::ncol(dsm)/2)
cell <- terra::cellFromRowCol(dsm, row, col)
xy <- terra::xyFromCell(dsm, cell)
viewpoint <- c(xy[,1], xy[,2])
# compute viewshed
# the height of the viewpoint is set as 6 feet
viewshed <- viewscape::compute_viewshed(dsm,
                                        viewpoint,
                                        offset_viewpoint = 6)
```

In real life, we can not see all theoretical visible objects at once with the limited view angle. To present a more realistic viewscape analysis, `sector_mask` is provided to extract

```{r eval = FALSE}
# subset the viewshed
sub_viewshed <- viewscape::sector_mask(viewshed, c(10,100))
```

The output 'viewshed' object can be written in raster format using `visualize_viewshed` for further analysis and visualization.

```{r eval = FALSE}
# create a raster of viewshed
viewshed_raster <- viewscape::visualize_viewshed(viewshed, outputtype = "raster")
# plot viewshed overlapping with the DSM
terra::plot(dsm, legend = FALSE)
terra::plot(viewshed_raster, add=TRUE, col = "red")
```

### 2. Compute visual magnitude
Binary viewshed only present the locations of visible contents, however, people may be also interested in quantifying the contribution of the visible content in a scene. Therefore, visual magnitude is used to estimate the proportion of each visible area unit within a viewshed. `visual_magnitude` function implements the visual magnitude algorithm for this purpose.

```{r eval = FALSE}
vm <- viewscape::visual_magnitude(viewshed, dsm)
```

### 3. Compute viewscape metrics
Viewscape metrics are meaningful in understanding the experience of urban landscapes, such as percieved restorativeness and visual walkability.  

