---
title: viewscape in R
author: Xiaohao Yang
date: '2024-02-13'
slug: viewscape
categories: ["tutorials"]
tags: ["R", "R package"]
---

This is an integrated tutorial about `viewscape` and `dsmSearch`. These two packages were initially one before getting published because I thought it would be cool if one can just specify a geographical area without DSM/DEM to compute a viewshed(s). This has been explained the purpose of making the `dsmSearch`, which was two functions for searching and downloading point cloud data via API.

Of course, `viewscape` is not designed for just computing the viewshed even if this is the core of the package. A set of function for calculating viewscape metrics ([see details](https://github.com/land-info-lab/viewscape/blob/master/README.md)) is the main contribution of this package. What's more, it provides the functionality to subset the viewshed based on the angle and orientation of the field of view and compute the [visual magnitude]() within the viewshed. 

Before I start, what is 'viewshed' and what is 'viewscape' by the way? Basically, I would say viewscape is a concept of analysis relying on viewshed, which is the visible area that can be seen from the  location of a observer. Viewscape analysis is to understand the composition of landscape and the structure of viewshed itself and visible objects within the viewshed.

```{r eval = FALSE}
# install packages
install.pakcages("viewscape")
install.pakcages("dsmSearch")
```

### 1.Compute viewshed
To compute a viewshed, a viewpoint and a DSM are essential. Let's download the LiDAR data from a small part of a city. In this tutorial, I take the Central Park (NYC) as an example. First, a bounding box (bbox) is used to locate an area to download the data. `lidR`, a powerful tool for processing LiDAR data, is used to generate DSM/DEM.

```{r eval = FALSE}
# bbox finder: http://bboxfinder.com/
bbox <- c(-73.976440,40.768793,-73.969628,40.773295)

# download LiDAR (.las) 
las <- dsmSearch::get_lidar(bbox = bbox, epsg = 2263)

# constrct DSM and DEM using .las data
dsm <- lidR::rasterize_canopy(las, 5, lidR::dsmtin())
dem <- lidR::rasterize_terrain(las, 5, lidR::tin())
```

In this case, I just set the center point of the DSM as the viewpoint. 

```{r eval = FALSE}
row <- trunc(terra::nrow(dsm)/2)
col <- trunc(terra::ncol(dsm)/2)
cell <- terra::cellFromRowCol(dsm, row, col)
xy <- terra::xyFromCell(dsm, cell)
viewpoint <- c(xy[,1], xy[,2])
# compute viewshed
# the height of the viewpoint is set as 6 feet
viewshed <- viewscape::compute_viewshed(dsm,
                                        viewpoint,
                                        offset_viewpoint = 6)
```

In real life, we can not see all theoretical visible objects at once with the limited view angle. To present a more realistic viewscape analysis, `sector_mask` is provided to extract

```{r eval = FALSE}
# subset the viewshed
sub_viewshed <- viewscape::sector_mask(viewshed, c(10,100))
```

The output 'viewshed' object can be written in raster format using `visualize_viewshed` for further analysis and visualization.

```{r eval = FALSE}
# create a raster of viewshed
viewshed_raster <- viewscape::visualize_viewshed(viewshed, outputtype = "raster")
# plot viewshed overlapping with the DSM
terra::plot(dsm, legend = FALSE)
terra::plot(viewshed_raster, add=TRUE, col = "red")
```

### 2. Compute visual magnitude
Binary viewshed only present the locations of visible contents, however, people may be also interested in quantifying the contribution of the visible content in a scene. Therefore, visual magnitude is used to estimate the proportion of each visible area unit within a viewshed. `visual_magnitude` function implements the visual magnitude algorithm for this purpose.

```{r eval = FALSE}
vm <- viewscape::visual_magnitude(viewshed, dsm)
```

### 3. Compute viewscape metrics for multiple viewpoints
Viewscape metrics are meaningful in understanding the experience of urban landscapes, such as perceived restorativeness and visual walkability. I used built-in data of this package to demonstrate whole process of computing viewscape metrics for multiple viewpoints.

First, I loaded in all data needed for the viewscape analysis. If you don't have DSM (or DTM), you can use your DTM (or DSM) as DSM (or DTM).

```{r}
# read a viewpoint
viewpoints <- sf::read_sf(system.file("test_viewpoints.shp", 
                                     package = "viewscape"))

# read DSM and DTM
dsm <- terra::rast(system.file("test_dsm.tif",
                               package ="viewscape"))
dtm <- terra::rast(system.file("test_dtm.tif", 
                               package ="viewscape"))

# read rasters of land use, canopy coverage, and building footprints
landuse <- terra::rast(system.file("test_landuse.tif",
                                   package ="viewscape"))
canopy <- terra::rast(system.file("test_canopy.tif", 
                                  package ="viewscape"))
building <- terra::rast(system.file("test_building.tif", 
                                    package ="viewscape"))
```

```{r figures-side, fig.show="hold", out.width="50%"}
terra::plot(dsm)
terra::plot(dtm)
terra::plot(landuse)
terra::plot(canopy)
terra::plot(building)
```

Second, I computed viewsheds for multiple viewpoints and specified `parallel` and `workers` in `compute_viewshed()`.

```{r}
# compute viewshed
viewsheds <- viewscape::compute_viewshed(dsm = dsm,
                                         viewpoints = viewpoints, 
                                         offset_viewpoint = 6)
```

Third, I used functions, including `calculate_diversity()`, `calculate_feature()`, and `calculate_viewmetrics()`, to calculate: 
- Shanon Diversity Index (SDI), 
- depth,
- vdepth,
- extent,
- horizontal,
- relief,
- skyline,
- percentage of canopy and building,
- Number of patches (Nump),
- Mean shape index (MSI),
- Edge density (ED),
- Patch size (PS),
- Patch density (PD)

Let's define a wrapped function. This has been saying that if one doesn't have some layers, for example canopy and buildings

```{r}
calculate <- function(viewsheds, dsm, dtm, land, canopy, building) {
  results <- matrix(0,length(viewsheds),16)
  colnames(results) <- c("x", "y", "Nump", "MSI", "ED", "PS", "PD",
                         "extent", "depth", "vdepth",
                         "horizontal", "relief", "skyline",
                         "sdi", "canopy", "building")
  masks <- list(canopy, building)
  for (i in 1:length(viewsheds)) {
    this_viewshed <- viewsheds[[i]]
    v_point <- this_viewshed@viewpoint
    metrics <- viewscape::calculate_viewmetrics(this_viewshed, 
                                                dsm, 
                                                dtm, 
                                                masks = masks)
    results[i,1] <- v_point[1] # x
    results[i,2] <- v_point[2] # y
    results[i,3] <- metrics$Nump # Nump
    results[i,4] <- metrics$MSI # MSI
    results[i,5] <- metrics$ED # ED
    results[i,6] <- metrics$PS # PS
    results[i,7] <- metrics$PD # PD
    results[i,8] <- metrics$extent # extent
    results[i,9] <- metrics$depth # depth
    results[i,10] <- metrics$vdepth # vdepth
    results[i,11] <- metrics$horizontal # horizontal
    results[i,12] <- metrics$relief # relief
    results[i,13] <- metrics$skyline # skyline
    results[i,14] <- viewscape::calculate_diversity(this_viewshed, land) # sdi
    results[i,15] <- viewscape::calculate_feature(this_viewshed, 
                                                  canopy, type = 2,
                                                  exclude_value = 0) # canopy
    results[i,16] <- viewscape::calculate_feature(this_viewshed, 
                                                  building, type = 2,
                                                  exclude_value = 0) # building
  }
  return(results)
} 
```

Now, run the function and check out the result.

```{r}
result <- calculate(viewsheds, dsm, dtm, landuse, canopy, building)
head(result)
```


